---
title: "WUP_04"
author: "Alexis Velasco"
date: "summer 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSScsv <- read.csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(GSScsv,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

The graph shows the difference in wordsum test scores between grade level.If you were to move the high school box it would not hit the graduate box and that is a visual for the large difference they have with one another. The lt high school box would not hit the junior college box either showing the difference between the two.


### Numerical Descriptive Results.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSScsv)
```

The mean of Lt high school is 7.75 where graduate students have a mean of 10.09 and the high school students with a mean of 8.28, shows that the box will not hit each other mostly likely giving up an extremely small chance for intersection between these five groups. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv)
summary.aov(model1)
```

 The chance of getting these results of having no difference between the five groups is 2.36 out of 10,000. This informs us that there is a difference in the test scores between the groups. With a low p value we are going to see what values differ from one another by using t tests. 
 
 
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 16.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```
When comparing all theses groups to  one another we can see that high school- graduate as well as LT high school- graduate confidence intervals do not cross into zero. The two not coming into contact with zero shows they have a difference.  
 
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The graph shows the difference in wordsum test scores between the grade levels.The chance of getting these results of having no difference between the five groups is 2.36 out of 10,000. This informs us that there is a large difference in the test scores between the groups.The mean of Lt high school is 7.75, where graduate students have a mean of 10.09 and the high school students with a mean of 8.28. Showing the boxes will not hit each other helped determined the small chance for intersection, so running a onewayanova and a tukey comparing each group to one another helps assist in the conclusion that the groups are have a difference between the test scores the the students grade level. 


